% !TEX TS-program = xelatex
\documentclass[11pt]{article}
\usepackage{xcolor}
\usepackage{lindrew}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{fontspec}
\pgfplotsset{compat=1.18}
\title{Math 4580: Abstract Algebra I}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=red,
    filecolor=red,
    urlcolor=red,
    citecolor=red
}
\author{Lecturer: \textbf{Professor Michael Lipnowski}\\Notes by: Farhan Sadeek}
\date{Spring 2025}

\begin{document}

\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{January 6, 2025}
We didn't have any, but Dr.\ Lipnowski did post a module on
\href{https://carmen.osu.edu}{{\texttt{carmen}}} about the syllabus and the
course. This semester we will be covering the first few chapters of the book
\textit{Abstract Algebra: Theory and Applications} by Thomas Judson. \\

\begin{definition}
    \vocab{Set}: A collection of distinct objects, considered as an object in its own right.\\
    \vocab{Axioms}: A collection of objects \( \mathrm{S} \) with assumed structural rules is defined by axioms.\\
    \vocab{Statement}: In logic or mathematics, an assertion that is either true or false.\\
    \vocab{Hypothesis and Conclusion}: In the statement ``If P, then Q'', P is the hypothesis and Q is the conclusion.\\
    \vocab{Mathematical Proof}: A logical argument that verifies the truth of a statement.\\
    \vocab{Proposition}: A statement that can be proven true.\\
    \vocab{Theorem}: A proposition of significant importance.\\
    \vocab{Lemma}: A supporting proposition used to prove a theorem or another proposition.\\
    \vocab{Corollary}: A proposition that follows directly from a theorem or proposition with minimal additional proof.
\end{definition}

\section{January 8, 2025}
Professor Lipnowski discussed Sam Lloyd's 15 puzzle. Each lecture will include
a mystery digit, contributing up to 5\% bonus to the final grade based on
correct guesses.

Certain course expectations:
\begin{itemize}
    \item All assignments (one every two weeks) and exams (one midterm and one final
          exam) will be take-home.
    \item All the problems from the course textbook.
    \item Collaboration is encouraged, but the work should be your own.
    \item For the exams, we are not supposed to talk to other friends.
\end{itemize}

\subsection{Functions}
\begin{definition}
    Let $A$ and $B$ be sets. A function $f: A \to B$ assigns exactly one output $f(a) \in B$ to every input $a \in A$.
\end{definition}
\begin{itemize}
    \item The set $A$ is called the \textbf{domain} of $f$.
    \item The set $B$ is called the \textbf{codomain} of $f$.
\end{itemize}

\begin{fact}The domain $A$, codomain $B$, and the assignment of outputs $f(a)$ to every input $a \in A$ are all part of the data defining a function. Just writing a formula like $f(x) = e^x$ does not determine a function, as the domain and codomain are not specified.
\end{fact}
For example:
\begin{itemize}
    \item $f: \mathbb{R} \to \mathbb{R}, f(x) = e^x$.
    \item $f: \mathbb{Q} \to \mathbb{Q}, f(x) = e^x$.
\end{itemize}
Although these functions use the same formula, their meanings are completely different because their domains and codomains differ.

\subsection{Graphs}
A function $f: A \to B$ is often identified with its \textbf{graph} in $A
    \times B$:
\[ \text{graph}(f) = \{(a, b) \in A \times B : b = f(a)\}. \]

\begin{lemma}
    Let $f: A \to B$ be a function. Its graph, $\text{graph}(f)$, passes the \textbf{vertical line test}: For every $a \in A$, $V_a := \{(a, b) \in A \times B : b \in B\}$ intersects $\text{graph}(f)$ in exactly one element.
\end{lemma}
\begin{center}
    \begin{tikzpicture}
        \draw[thick,->] (-2,0) -- (2,0) node[right] {$x$};
        \draw[thick,->] (0,-2) -- (0,2) node[above] {$y$};
        \draw[domain=-1.5:1.5,smooth,variable=\x,blue,thick] plot ({\x},{\x^2});
        \draw[dashed] (-1,0) -- (-1,1) node[midway,left] {Vertical line test};
        \draw[fill=red] (-1,1) circle (2pt);
    \end{tikzpicture}
\end{center}

\begin{proposition}
    Let $G \subseteq A \times B$ be any subset passing the vertical line test, i.e., for all $a \in A$, $V_a \cap G$ consists of exactly one element. Then $G = \text{graph}(f)$ for a unique function $f: A \to B$.
\end{proposition}
\begin{proof}
    If $G = \{(a, b) \mid b \in B\}$ satisfies the vertical line test, define $f: A \to B$ by $f(a) = b$. Then $G = \text{graph}(f)$.
\end{proof}

\begin{definition}
    A subset $R \subseteq A \times B$ is called a \textbf{relation}. The vertical line test distinguishes graphs of functions from more general relations.
\end{definition}
\subsection{Examples}
\begin{itemize}
    \item Let $S = \{(x, y) \in \mathbb{R}^2 : x^2 + y^2 = 1\}$ (the unit circle). This
          is a relation but not the graph of a function because it fails the vertical
          line test: The vertical line $x = 0$ intersects the circle at two points.
    \item Visual depiction of a unit circle:
          \begin{center}
              \begin{tikzpicture}
                  \draw[thick,->] (-2,0) -- (2,0) node[right] {$x$};
                  \draw[thick,->] (0,-2) -- (0,2) node[above] {$y$};
                  \draw[thick] (0,0) circle (1.5);
                  \node[below right] at (2,2.5) {$x^2 + y^2 = 1$};
                  \draw[dashed] (0.5,-1.5) -- (0.5,1.5);
              \end{tikzpicture}
          \end{center}
    \item Let $A = \{1, 2, 3\}$, $B = \{4, 5\}$. The number of functions from $A$ to $B$
          is $2^3 = 8$, corresponding to the $8$ associated graphs in $A \times B$.
    \item The number of relations from $A$ to $B$ is $2^{ |a| \cdot  |b|} = 2^{3 \cdot 2} =
              64$, containing the $8$ graphs of functions from $A$ to $B$.
\end{itemize}

\begin{fact}
    The notion of relation is much more permissive than the notion of functions.
\end{fact}

\subsection{Visualizing Functions as Directed Edges}
A function $f: A \to B$ can be visualized as a collection of directed edges
$(a, f(a)) \in A \times B$. Each element of $A$ has exactly one outgoing edge
in the graph.

\begin{center}
    \begin{tikzpicture}
        \node[circle,draw] (A1) at (0,1) {$1$};
        \node[circle,draw] (A2) at (0,0) {$2$};
        \node[circle,draw] (A3) at (0,-1) {$3$};
        \node[circle,draw] (B1) at (2,0.5) {$4$};
        \node[circle,draw] (B2) at (2,-0.5) {$5$};
        \draw[->] (A1) -- (B1);
        \draw[->] (A2) -- (B2);
        \draw[->] (A3) -- (B1);
    \end{tikzpicture}
\end{center}
\section{January 10, 2025}
\subsection{More about function}

\(f : A \rightarrow B\) is a function.
\begin{itemize}
    \item injective (one-to-one)
    \item surjective (onto)
    \item bijective (one-to-one and onto)
\end{itemize}

\begin{definition}
    Let \(f: A \rightarrow B\) be a function. \(f\) is injective if for all \(x, y \in A\), if \(f(x) = f(y)\) then \(x = y\).
\end{definition}
\begin{example}
    Consider the function $f: \{1, 2, 3\} \to \{a, b, c\}$ defined by the following assignments:
    \[
        f(1) = a, \quad f(2) = b, \quad f(3) = c
    \]
    This function is bijective because it is both injective and surjective.
\end{example}

\begin{example}
    Consider the function $f: \mathbb{R} \to \mathbb{R}$ defined by $f(x) = 2x + 3$. We claim that $f$ is injective.

    To prove this, suppose $f(a) = f(b)$ for some $a, b \in \mathbb{R}$. Then:
    \[
        2a + 3 = 2b + 3
    \]
    Subtracting 3 from both sides, we get:
    \[
        2a = 2b
    \]
    Dividing both sides by 2, we obtain:
    \[
        a = b
    \]
    Therefore, $f$ is injective.
\end{example}

\begin{example}
    Consider the function $f: \mathbb{R} \to \mathbb{R}$ defined by $f(x) = x^2$. We claim that $f$ is not injective.

    To see this, observe that $f(2) = 4$ and $f(-2) = 4$. Since $f(2) = f(-2)$ but
    $2 \neq -2$, the function $f$ is not injective.
\end{example}

\begin{example}
    Consider the sets $A = \{1, 2, 3\}$ and $B = \{4, 5\}$. Define the function $f: A \to B$ by the following assignments:
    \[
        f(1) = 4, \quad f(2) = 5, \quad f(3) = 4
    \]
    This function is not injective because $f(1) = f(3) = 4$ but $1 \neq 3$.
\end{example}
Here we have two elements in set $B$ but there are three elements in input in set $A$. That's why there has to be a collision. Since $ |a| >  |b|$, there must be a collision. If $f(1), f(2), f(3) \in B$, and $ |b| = 2$ Tow of those must be the same. That means that f is not one-to-one.

\begin{definition}
    A function $f: A \to B$ is called \textbf{surjective} (or \textbf{onto}) if for every element $b \in B$, there exists at least one element $a \in A$ such that $f(a) = b$.
\end{definition}
\begin{example}
    Consider the function $f: \mathbb{R} \to \mathbb{R}$ defined by $f(x) = x$. This function is surjective because for every $y \in \mathbb{R}$, we can find an $x \in \mathbb{R}$ such that $f(x) = y$ (specifically, $x = y$).
\end{example}
\begin{example}
    Consider the function $f: \mathbb{R} \to \mathbb{R}$ defined by $f(x) = x^2$. This function is not surjective because there is no $x \in \mathbb{R}$ such that $f(x) = -1$.
\end{example}
\begin{example}
    Consider the sets $A = \{1, 2, 3\}$ and $B = \{4, 5\}$. Define the function $f: A \to B$ by the following assignments:
    \[
        f(1) = 4, \quad f(2) = 4, \quad f(3) = 4
    \]
    This function is surjective because every element in $B$ is mapped to by at
    least one element in $A$. However, it is not injective because $f(1) = f(2) =
        f(3) = 4$ but $1 \neq 2 \neq 3$.
\end{example}

\begin{example}
    Consider the sets $A = \{1, 2, 3\}$ and $B = \{4, 5\}$. Define the function $f: A \to B$ by the following assignments:
    \[
        f(1) = 4, \quad f(2) = 4, \quad f(3) = 5
    \]
    This function is surjective because every element in $B$ is mapped to by at
    least one element in $A$. However, it is not injective because $f(1) = f(2) =
        4$ but $1 \neq 2$.
\end{example}

\begin{example}
    Consider the sets $A = \{1, 2, 3\}$ and $B = \{4, 5, 6, 7\}$. Define the function $f: A \to B$ by the following assignments:
    \[
        f(1) = 4, \quad f(2) = 5, \quad f(3) = 6
    \]
    This function is neither injective nor surjective. It is not injective because
    $ |a| <  |b|$, and it is not surjective because the element $7 \in B$ is not
    mapped to by any element in $A$.
\end{example}

\begin{definition}
    A function $f: A \to B$ is called \textbf{range} if it is the set of all possible outputs of $f$. Formally, the range of $f$ is defined as:
    \[
        \text{range}(f) = \{ f(a) \mid a \in A \}
    \]
\end{definition}
\begin{fact}
    The range of a function $f: A \to B$ is a subset of the codomain $B$. A function is surjective if and only if its range is equal to its codomain.
\end{fact}
\begin{example}
    Consider the function $f: \{1, 2, 3\} \to \{4, 5\}$ defined by the following assignments:
    \[
        f(1) = 4, \quad f(2) = 5, \quad f(3) = 4
    \]
    The range of $f$ is $\{4, 5\}$, which is equal to the codomain $\{4, 5\}$.
    Therefore, $f$ is surjective.
\end{example}
\begin{example}
    Consider the function $f: \{1, 2, 3\} \to \{4, 5, 6\}$ defined by the following assignments:
    \[
        f(1) = 4, \quad f(2) = 5, \quad f(3) = 4
    \]
    The range of $f$ is $\{4, 5\}$, which is a subset of the codomain $\{4, 5, 6\}$
    but not equal to it. Therefore, $f$ is not surjective.
\end{example}
\begin{fact}
    f is \vocab{surjective} means that the \vocab{range} of $f$ is equal to the \vocab{codomain} of $f$.
\end{fact}

\begin{definition}
    A function $f: A \to B$ is called \textbf{bijective} (or a \textbf{bijection}) if it is both injective and surjective or (one-to-one correspondence). This means that every element in $B$ is mapped to by exactly one element in $A$.
\end{definition}
\begin{example}
    Consider the function $f: \{1, 2, 3\} \to \{4, 5, 6\}$ defined by the following assignments:
    \[
        f(1) = 5, \quad f(2) = 4, \quad f(3) = 6
    \]
    This function is bijective because it is both injective (no two elements in $A$
    map to the same element in $B$) and surjective (every element in $B$ is mapped
    to by some element in $A$).
\end{example}
Here we can notice that we have a matching between the elements of $A$ and $B$, every element of set $A$ has one and only mapping into the set $B$.

\begin{example}
    Consider the function $f: \mathbb{N} \to \mathbb{Z}$ defined by $f(x) = x$, where $\mathbb{N}$ is the set of natural numbers and $\mathbb{Z}$ is the set of integers. We claim that $f$ is injective but not surjective.

    To prove that $f$ is injective, suppose $f(a) = f(b)$ for some $a, b \in \mathbb{N}$. Then:
    \[
        a = b
    \]
    Therefore, $f$ is injective because no two different elements in $\mathbb{N}$ map to the same element in $\mathbb{Z}$.

    However, $f$ is not surjective because there are elements in $\mathbb{Z}$ that are not in the range of $f$. For example, there is no $x \in \mathbb{N}$ such that $f(x) = -1$. Therefore, $f$ is not surjective.
\end{example}


\begin{example}
    Consider the function $f: \mathbb{Z} \to \mathbb{N}$ defined by:
    \[
        f(n) = 
        \begin{cases} 
            2n & \text{if } n \geq 0 \\
            2|n| - 1 & \text{if } n < 0 
        \end{cases}
    \]
    We claim that $f$ is bijective.

    To prove that $f$ is injective, suppose $f(a) = f(b)$ for some $a, b \in \mathbb{Z}$. We need to show that $a = b$.
    \begin{itemize}
        \item If $a \geq 0$ and $b \geq 0$, then $f(a) = 2a$ and $f(b) = 2b$. Since $f(a) = f(b)$, we have $2a = 2b$, which implies $a = b$.
        \item If $a < 0$ and $b < 0$, then $f(a) = 2 |a| - 1$ and $f(b) = 2 |b| - 1$. Since $f(a) = f(b)$, we have $2 |a| - 1 = 2 |b| - 1$, which implies $ |a| =  |b|$ and hence $a = b$.
        \item If $a \geq 0$ and $b < 0$, then $f(a) = 2a$ and $f(b) = 2 |b| - 1$. Since $f(a) = f(b)$, we have $2a = 2  |b| - 1$, which is a contradiction because $2a$ is even and $2 |b| - 1$ is odd. Therefore, this case cannot occur.
        \item If $a < 0$ and $b \geq 0$, then $f(a) = 2 |a| - 1$ and $f(b) = 2b$. Since $f(a) = f(b)$, we have $2 |a| - 1 = 2b$, which is a contradiction because $2 |a| - 1$ is odd and $2b$ is even. Therefore, this case cannot occur.
    \end{itemize}
    Therefore, $f$ is injective.

    To prove that $f$ is surjective, let $m \in \mathbb{N}$. We need to find $n \in \mathbb{Z}$ such that $f(n) = m$.
    \begin{itemize}
        \item If $m$ is even, say $m = 2k$ for some $k \in \mathbb{N}$, then $f(k) = 2k = m$.
        \item If $m$ is odd, say $m = 2k + 1$ for some $k \in \mathbb{N}$, then $f(-(k+1)) = 2|-(k+1)| - 1 = 2(k+1) - 1 = m$.
    \end{itemize}
    Therefore, $f$ is surjective.

    Since $f$ is both injective and surjective, it is bijective.
\end{example}


\begin{example}
    Consider the function $f: \mathbb{Z} \to \mathbb{N}$ defined by:
    \[
        f(n) = 
        \begin{cases} 
            2n & \text{if } n \geq 0 \\
            2(-n - 1) + 1 & \text{if } n < 0 
        \end{cases}
    \]
    We claim that $f$ is bijective.
\end{example}


%\section{January 13, 2025}
%\section{January 15, 2025}
%\section{January 17, 2025}
%\section{January 20, 2025}
%\section{January 22, 2025}
%\section{January 24, 2025}
%\section{January 27, 2025}
%\section{January 29, 2025}
%\section{January 31, 2025}
%\section{February 3, 2025}
%\section{February 5, 2025}
%\section{February 7, 2025}
%\section{February 10, 2025}
%\section{February 12, 2025}
%\section{February 14, 2025}
%\section{February 17, 2025}
%\section{February 19, 2025}
%\section{February 21, 2025}
%\section{February 24, 2025}
%\section{February 26, 2025}
%\section{February 28, 2025}
%\section{March 3, 2025}
%\section{March 5, 2025}
%\section{March 7, 2025}
%\section{March 17, 2025}
%\section{March 19, 2025}
%\section{March 21, 2025}
%\section{March 24, 2025}
%\section{March 26, 2025}
%\section{March 28, 2025}
%\section{March 31, 2025}
%\section{April 2, 2025}
%\section{April 4, 2025}
%\section{April 7, 2025}
%\section{April 9, 2025}
%\section{April 11, 2025}
%\section{April 14, 2025}
%\section{April 16, 2025}
%\section{April 18, 2025}
%\section{April 21, 2025}
%\section{April 23, 2025}
%\section{April 25, 2025}
%\section{April 28, 2025}

\end{document}
