\documentclass[11pt]{article}
%\documentclass[11pt]{article}
\usepackage{lindrew}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{fontspec}
\title{Physics 2301: Intermediate Mechanics II}
\author{Lecturer: \textbf{Professor Antonio Boveia}\\Homework by: Farhan Sadeek}
\date{Spring 2025}

\begin{document}
\maketitle
\section*{Problem 1}
\begin{itemize}
    \item \textbf{What's your major / what kind of job would you like after graduation?} \\
          I am currently a senior in high school, and I am undecided about my major. I am taking this class because I would like to expand my knowledge in physics, and as a preparation for my USA Physics Olympiad (USAPhO). My parents can't afford to pay for tutoring or courses, so I decided to self-teach myself physics using books and this course (paid by the school district) will help me both in terms of knowledge and preparation for the olympiad. Once I grow up, I would like to see myself as a quantum physicist or a particle physicist. \\

    \item \textbf{What would you like to know about physics that isn't taught in a class?} \\
          Most likely how the theory of relativity is useful in our daily lives. I guess we sort of see its usage in telecommunications. But I would really like to know how these theoretical concepts could be used in real life. \\

    \item \textbf{How do you feel about your math preparation for Physics 2300?} \\
          I am feeling somewhat confident because I took \textbf{Math 2153, Math 2568, and Math 4512} but I am a little rusty on \textbf{Math 2568} so I would like to review that this and the following weeks before physics starts to consume a lot of my time. :'(\\

    \item \textbf{Do you like the limericks in the textbook?} \\
          Yeah absolutely, they sometimes make me wonder if Dr. Morin is a poet or a physicist. \\

    \item \textbf{What was your favorite part of last semester?} \\
          \textbf{Physics 2300} was definitely very challenging but I met some lifelong friends there and understood that I need both hard and smart work to succeed in this class. \\

    \item \textbf{Are you a dog or cat person? (Or eww, why would anyone want a pet?)} \\
          Either is fine with me, but I never had a pet as of now, but I would like to get either a dog or a cat in the future. \\
\end{itemize}

\section*{Problem 2}
\textbf{What would you change about the “flipped classroom” format?}
\\ I wouldn't change anything. It would be really helpful if you could suggest some \textbf{optional} \href{https://youtube.com}{YouTube} videos that we could watch along with your lectures, so that it's easier for me to catch up. I am personally a more visual learner, and I think there are perhaps people like me too, so that's a request not a requirement.
\end{document}
